\textbf{\supplement}\vskip 2ex \mytitle\vskip 2ex

### 2023-01-31 generated/legend.tex
Updating to January 2023 submission to Computer Journal

\PassOptionsToPackage{hyphens}{url}
\documentclass{article}
\renewcommand\thesubsection{\thesection.\alph{subsection}}
%\usepackage{hyperref}
%\usepackage{refcount}
\usepackage{url}
\usepackage{doi}
\usepackage{longtable}
\usepackage{color}
\usepackage{amsmath,amssymb}
\def\httpURL#1{\href{http://#1}{\textcolor{blue}{#1}}}
\def\LONGhttpURL#1#2{\href{http://#1}{\textcolor{blue}{#2}}}

% expand paper-seb-macros.tex
% These macros allow the paper and the supplementary material to share a single sequence of bibliography citations

\DeclareUrlCommand\doi{\def\UrlLeft{{\textrm{DOI}}~}}
\DeclareUrlCommand\url{\def\UrlLeft{{\textrm{URL}}~}}

\renewcommand{\thefootnote}{\fnsymbol{footnote}}

\def\mytitle{Computational science: \hbox{A (fixable) failure of software~engineering}}
%Achieving professional software engineering \hbox{in scientific research}}

\def \citeeg#1{(e.g., \cite{#1})}

\def\inputifexists#1#2{\IfFileExists{#1}{\input{#1}}{\typeout{No file #1. #2}}}

% calculate percentages
\newcount \t
\newcount \tu
\def\pc#1#2{\t = #1%
\multiply \t by 100000%
\divide \t by #2% ?? 0 times percentage
\tu = \t
\divide \t by 1000% percentage, but no idea of remainder
\multiply \t by 1000% 1000 times percentage, with 00 as bottom digits
\advance \tu by -\t% bottom two digits
\divide \t by 1000\relax%
% round to nearest integer
% if ending in > .5 round up
% if ending in exactly .5 round towards nearest even number
\ifnum \tu > 500 % round up
	\advance \t by 1 
\else
	\ifnum \tu = 500 
		\ifodd \t % it's odd, so round up to even number 
			\advance \t by 1
		\else % leave rounded to even number 	
		\fi
	\fi
\fi
\the\t\%}

% insert commas into large numbers
\def\commarise#1{%[#1]
	\t=#1
    \divide \t by 1000
    \ifnum \t > 0
    	\the\t,%
    	\tu=#1
    	\multiply \t by 1000
    	\advance \tu by -\t
		\ifnum \tu < 100 0\fi
		\ifnum \tu < 10 0\fi
		\the\tu
    \else
    	#1%
    \fi
}

% simple macro to use a number register to pluralise (add an s) to words....
% \plural{\t}{fred} => \the\t\ freds if plural, or => one fred if singular
\def\plural#1#2{\ifnum #1=1 
	one #2%
\else
	\the#1\ #2s%
\fi}

% initialisBibliography <title> <starting number> <introductory text>
\newcount\bibciten \bibciten=0
\def\initialiseBibliography#1#2#3{%
	\global\def\refname{#1}
    \global\bibciten=#2
    \global\def\startBibliography{#3}
}

\def\bibskip{\vskip 1ex} % the gap between bib items generated by data.js

% read in any constants defined from the JSON data
% expand generated-constants.tex
\def\journalBreakdown{\emph{Lancet Digital Health\/} ($N=6$), \emph{Nature Digital Medicine\/} ($N=12$) and \emph{Royal Society Open Science\/} ($N=14$)}
\def\tabularJournalBreakdown{&\hbox to 3em {\hfill 6}\hskip 1em \emph{Lancet Digital Health}\\&\hbox to 3em {\hfill 12}\hskip 1em \emph{Nature Digital Medicine}\\&\hbox to 3em {\hfill 14}\hskip 1em \emph{Royal Society Open Science}\\}
\global\newcount \dataN \global\dataN=32
\global\newcount \countAuthors \global\countAuthors=264
\global\newcount \countHasBreach \global\countHasBreach=11
\global\newcount \countHasPolicy \global\countHasPolicy=26
\global\newcount \countUsesVersionControlRepository \global\countUsesVersionControlRepository=10
\global\newcount \counthasDataRepository \global\counthasDataRepository=9
\global\newcount \countNoCodeInRepo \global\countNoCodeInRepo=1
\global\newcount \numberOfJournals \global\numberOfJournals=3
\global\newcount \countCodetested \global\countCodetested=0
\global\newcount \hasDevelopedRigorously \global\hasDevelopedRigorously=0
% end expanding generated-constants.tex
% expand generated-info-for-main.tex
\newlabel{supplementary-material-makefiles}{{2}{11}{DUMMYA}{DUMMYB}{DUMMYC}}
\newlabel{on-code-data-publication}{{3}{14}{XXX}{YYYY}{ZZZZ}}
\newlabel{supplementary-Speigelhalter-section}{{4}{17}{DUMMYA}{DUMMYB}{DUMMYC}}
\def\MaxMainPaperCitationNumber{64}
\newlabel{supplementary-journal-policies-section}{{5}{25}{DUMMYA}{DUMMYB}{DUMMYC}}
\newlabel{supplementary-summary-table}{{5}{30}{DUMMY1}{DUMMY2}{DUMMY3}}
\expandafter\def \csname cite-MetricSelectionFramework\endcsname{[97]}
\expandafter\def \csname cite-PENet\endcsname{[101]}
\expandafter\def \csname cite-PostoperativeOutcomes.RiskNet\endcsname{[103]}
\expandafter\def \csname cite-philter-ucsf\endcsname{[104]}
\expandafter\def \csname cite-AI-CDSS-Cardiovascular-Silo\endcsname{[105]}
\expandafter\def \csname cite-SiameseChange\endcsname{[107]}
\expandafter\def \csname cite-dryad.1g1jwstrw\endcsname{[108]}
\expandafter\def \csname cite-rsos.192210\endcsname{[109]}
\expandafter\def \csname cite-ichHKrWj7hqlznOaR6NQVzITgp40dlqWvWAgAxyafiQ\endcsname{[110]}
\expandafter\def \csname cite-rsos.200566\endcsname{[111]}
\expandafter\def \csname cite-?view.only=87ae173f775b40d79d6cd0fdcf6d4a9c\endcsname{[112]}
\expandafter\def \csname cite-dryad.vx0k6djnr\endcsname{[115]}
\expandafter\def \csname cite-lactModel\endcsname{[116]}
\expandafter\def \csname cite-LRM\endcsname{[118]}
\expandafter\def \csname cite-1\endcsname{[119]}
\expandafter\def \csname cite-manifold-ga\endcsname{[124]}
\expandafter\def \csname cite-blast-ct\endcsname{[126]}
% end expanding generated-info-for-main.tex

\def\supplement{Supplemental Material}
% end expanding paper-seb-macros.tex

\makeatletter
      \renewcommand{\thefootnote}{\arabic{footnote}}
      {\let\oldbibcite=\bibcite
      % example: \bibcite{bad-code}{{31}{2020}{{Richards and Boudnik}}{{Richards and Boudnik}}}
      \newcount\mainNumberOfReferences \mainNumberOfReferences=0
      \def\bibcite#1#2{\global\advance\mainNumberOfReferences by 1\oldbibcite{#1}{#2}}
      % expand paper-seb-main.aux
\relax 
\providecommand\hyper@newdestlabel[2]{}
\providecommand\HyperFirstAtBeginDocument{\AtBeginDocument}
\HyperFirstAtBeginDocument{\ifx\hyper@anchor\@undefined
\global\let\oldcontentsline\contentsline
\gdef\contentsline#1#2#3#4{\oldcontentsline{#1}{#2}{#3}}
\global\let\oldnewlabel\newlabel
\gdef\newlabel#1#2{\newlabelxx{#1}#2}
\gdef\newlabelxx#1#2#3#4#5#6{\oldnewlabel{#1}{{#2}{#3}}}
\AtEndDocument{\ifx\hyper@anchor\@undefined
\let\contentsline\oldcontentsline
\let\newlabel\oldnewlabel
\fi}
\fi}
\global\let\hyper@last\relax 
\gdef\HyperFirstAtBeginDocument#1{#1}
\providecommand*\HyPL@Entry[1]{}
\bibstyle{plos2015}
\citation{abelson}
\HyPL@Entry{0<</S/D>>}
\citation{fixit}
\citation{example-stats}
\citation{Speigelhalter}
\@writefile{toc}{\contentsline {section}{\numberline {1}Introduction}{2}{section.1}\protected@file@percent }
\newlabel{problems}{{1}{2}{Introduction}{section.1}{}}
\citation{Ben,se-bias}
\citation{whitty}
\citation{assurance-case}
\@writefile{toc}{\contentsline {subsection}{\numberline {(1.a)}The role of code in science and scientific publication}{4}{subsection.1.1}\protected@file@percent }
\citation{assessing-quality}
\citation{critiques,diagnosis-reviews}
\citation{knuth}
\citation{science-review}
\citation{diagnosis-reviews}
\citation{tripod}
\citation{diagnosis-reviews}
\@writefile{toc}{\contentsline {subsection}{\numberline {(1.b)}Bugs, code and programming}{5}{subsection.1.2}\protected@file@percent }
\newlabel{knowledge}{{(1.b)}{5}{Bugs, code and programming}{subsection.1.2}{}}
\@writefile{toc}{\contentsline {section}{\numberline {2}State of the art in pandemic modeling}{5}{section.2}\protected@file@percent }
\citation{nature-summary,ICmodel}
\citation{avianFluModel,originalICmodel}
\citation{ICmodel,avianFluModel,originalICmodel}
\citation{tweet}
\citation{ferguson-interview}
\citation{basic-reproducibilty}
\citation{thumbs-up}
\citation{nature-summary,ICmodel}
\citation{codecheck,thumbs-up}
\citation{NVP}
\citation{nvp-ferguson}
\citation{refactoring}
\citation{bad-code}
\citation{pseudo}
\citation{relit}
\citation{nature-review}
\citation{nature-review}
\citation{psychological-modeling}
\citation{psychological-modeling}
\citation{actr}
\@writefile{toc}{\contentsline {subsection}{\numberline {(2.a)}Science beyond epidemiology}{7}{subsection.2.1}\protected@file@percent }
\citation{psychological-modeling}
\citation{Ben,nature-review,se-bias}
\citation{ourReview}
\citation{hamming}
\citation{NVP}
\citation{unfinished}
\citation{unfinished,lancet-unfinished}
\citation{hoare}
\citation{notebooks,popper}
\citation{machine-learning-reproducibility}
\citation{relit}
\citation{ICmodel}
\citation{ICmodel}
\citation{ICmodel}
\citation{flu-model}
\@writefile{lot}{\contentsline {table}{\numberline {1}{\ignorespaces Overview of peer-reviewed paper sample.\relax }}{9}{table.caption.4}\protected@file@percent }
\providecommand*\caption@xref[2]{\@setref\relax\@undefined{#1}}
\newlabel{table-overview}{{1}{9}{Overview of peer-reviewed paper sample.\relax }{table.caption.4}{}}
\@writefile{toc}{\contentsline {section}{\numberline {3}A pilot survey of peer-reviewed research relying on code}{9}{section.3}\protected@file@percent }
\newlabel{survey-section}{{3}{9}{A pilot survey of peer-reviewed research relying on code}{section.3}{}}
\@writefile{toc}{\contentsline {subsection}{\numberline {(3.a)}Methodology}{9}{subsection.3.1}\protected@file@percent }
\@writefile{lot}{\contentsline {table}{\numberline {2}{\ignorespaces Summary of survey results.\relax }}{10}{table.caption.5}\protected@file@percent }
\newlabel{table-summary}{{2}{10}{Summary of survey results.\relax }{table.caption.5}{}}
\@writefile{lot}{\contentsline {table}{\numberline {3}{\ignorespaces Sizes of repositories, with approximate sizes of code (in kLOC) and data for all available GitHub repositories reviewed in the survey, plus \texttt  {covid-sim} \cite  {ICmodel} for comparison. Sizes are approximate because in all repositories code and data are conceptually interchangeable (an issue explained in the Supplemental Material), so choices were made in the survey to avoid double-counting. Many repositories rely on downloading additional code and data, which is not counted in the table as the additional required material is not in the repository cited in the paper. Paper [103]\ has nothing in its repository except a single file still saying ``Code coming soon...'' despite, at the time of cloning and checking repositories, 21\ months had already elapsed since the paper referring to the code was published.\relax }}{11}{table.caption.6}\protected@file@percent }
\newlabel{table-repo-summary}{{3}{11}{Sizes of repositories, with approximate sizes of code (in kLOC) and data for all available GitHub repositories reviewed in the survey, plus \texttt {covid-sim} \cite {ICmodel} for comparison. Sizes are approximate because in all repositories code and data are conceptually interchangeable (an issue explained in the \supplement ), so choices were made in the survey to avoid double-counting. Many repositories rely on downloading additional code and data, which is not counted in the table as the additional required material is not in the repository cited in the paper. Paper \csname cite-PostoperativeOutcomes.RiskNet\endcsname \ has nothing in its repository except a single file still saying ``Code coming soon...'' despite, at the time of cloning and checking repositories, \the \pubdelayinmonths \ months had already elapsed since the paper referring to the code was published.\relax }{table.caption.6}{}}
\citation{open-source}
\citation{stability}
\@writefile{toc}{\contentsline {subsection}{\numberline {(3.b)}Summary of results}{12}{subsection.3.2}\protected@file@percent }
\citation{notebooks}
\citation{cbc}
\@writefile{toc}{\contentsline {section}{\numberline {4}Discussion}{13}{section.4}\protected@file@percent }
\@writefile{toc}{\contentsline {subsection}{\numberline {(4.a)}A call to action}{13}{subsection.4.1}\protected@file@percent }
\newlabel{summary}{{(4.a)}{13}{A call to action}{subsection.4.1}{}}
\citation{ethics-code}
\citation{acm-artifacts}
\citation{redmill,iec61508}
\citation{basic-reproducibilty,open-source}
\citation{ABCs-SE}
\citation{lancet-retracted}
\citation{science-lancet1,science-lancet2}
\citation{lancet-learning}
\@writefile{toc}{\contentsline {subsection}{\numberline {(4.b)}Suggestions for further work}{15}{subsection.4.2}\protected@file@percent }
\newlabel{jvs-policy}{{(4.b)}{15}{Suggestions for further work}{subsection.4.2}{}}
\citation{jvs1}
\citation{jvs2,jvs3}
\citation{jvs1}
\citation{numerals}
\citation{fda}
\citation{tweet,ferguson-interview}
\citation{excel-fiasco}
\citation{science-delays}
\citation{whitty}
\@writefile{toc}{\contentsline {section}{\numberline {5}Conclusions}{16}{section.5}\protected@file@percent }
\citation{fixit}
\citation{reproducibility,relit,popper}
\citation{parliamentary-evidence,my-parliamentary-evidence}
\bibdata{paper-seb-main.bib}
\bibcite{abelson}{1}
\bibcite{fixit}{2}
\bibcite{example-stats}{3}
\bibcite{Speigelhalter}{4}
\bibcite{Ben}{5}
\bibcite{se-bias}{6}
\bibcite{whitty}{7}
\bibcite{assurance-case}{8}
\bibcite{assessing-quality}{9}
\citation{assessing-quality-conference}
\bibcite{critiques}{10}
\bibcite{diagnosis-reviews}{11}
\bibcite{knuth}{12}
\bibcite{science-review}{13}
\bibcite{tripod}{14}
\bibcite{nature-summary}{15}
\bibcite{ICmodel}{16}
\bibcite{avianFluModel}{17}
\bibcite{originalICmodel}{18}
\bibcite{tweet}{19}
\bibcite{ferguson-interview}{20}
\bibcite{basic-reproducibilty}{21}
\bibcite{thumbs-up}{22}
\bibcite{codecheck}{23}
\bibcite{NVP}{24}
\bibcite{nvp-ferguson}{25}
\bibcite{refactoring}{26}
\bibcite{bad-code}{27}
\bibcite{pseudo}{28}
\bibcite{relit}{29}
\bibcite{nature-review}{30}
\bibcite{psychological-modeling}{31}
\bibcite{actr}{32}
\bibcite{ourReview}{33}
\bibcite{hamming}{34}
\bibcite{unfinished}{35}
\bibcite{lancet-unfinished}{36}
\bibcite{hoare}{37}
\bibcite{notebooks}{38}
\bibcite{popper}{39}
\bibcite{machine-learning-reproducibility}{40}
\bibcite{flu-model}{41}
\bibcite{open-source}{42}
\bibcite{stability}{43}
\bibcite{cbc}{44}
\bibcite{ethics-code}{45}
\bibcite{acm-artifacts}{46}
\bibcite{redmill}{47}
\bibcite{iec61508}{48}
\bibcite{ABCs-SE}{49}
\bibcite{lancet-retracted}{50}
\bibcite{science-lancet1}{51}
\bibcite{science-lancet2}{52}
\bibcite{lancet-learning}{53}
\bibcite{jvs1}{54}
\bibcite{jvs2}{55}
\bibcite{jvs3}{56}
\bibcite{numerals}{57}
\bibcite{fda}{58}
\bibcite{excel-fiasco}{59}
\bibcite{science-delays}{60}
\bibcite{reproducibility}{61}
\bibcite{parliamentary-evidence}{62}
\bibcite{my-parliamentary-evidence}{63}
\citation{parliamentary-evidence}
\bibcite{assessing-quality-conference}{64}
\newlabel{LastPage}{{}{22}{}{page.22}{}}
\xdef\lastpage@lastpage{22}
\xdef\lastpage@lastpageHy{22}
\gdef \@abspage@last{22}
% end expanding paper-seb-main.aux
      }
\makeatother

\begin{document}

\title{\textbf{\supplement}\vskip 2ex \mytitle\vskip 2ex}

\author{Harold Thimbleby, \texttt{harold@thimbleby.net}}

%\orcid{0000-0003-2222-4243}

% To make various LaTeX processors do the right thing with page size.
\def\pprw{8.5in}
\def\pprh{11in}
\special{papersize=\pprw,\pprh}
\setlength{\paperwidth}{\pprw}
\setlength{\paperheight}{\pprh}
\setlength{\pdfpagewidth}{\pprw}
\setlength{\pdfpageheight}{\pprh}

\def\href#1{#1}

\maketitle

% this will be used to save cross references that can be used in the main paper too
\newwrite\infofile
\immediate\openout\infofile=generated-info-for-main.tex

\def\dotalph#1{\ifnum\csname c@#1\endcsname>0 .\alph{#1}\fi}

% writing to the aux file, this file generates labels like
% \newlabel{supplementary-best-practice}{{2}{5}{Software engineering best practice}{section.2}{}}
% whereas the main file using IEEEtrans class generates
% \newlabel{table-summary}{{I}{5}} or \newlabel{jvs-policy}{{V}{4}} etc

\makeatletter
\def\globalLabel#1{% we also need journal code-policies to be used in the main paper...
\immediate\write\infofile{\bslash newlabel{#1}{{\arabic{section}}{\thepage}{DUMMYA}{DUMMYB}{DUMMYC}}}
% as well as in the current document....
\immediate\write\@auxout{\string\newlabel{#1}{{\arabic{section}\dotalph{subsection}}{\thepage}{AAA}{BBB}{CCC}}}
}
\makeatother

% start supplementary material's page counters to continue on from the paper's page numbering
%\setcounterpageref{page}{LastPage}
%\addtocounter{page}{1} % otherwise we start with the same page that the main paper ended on (we want to be on the next page:-)!

\setcounter{section}{0}
\def\n#1{{\sf\bfseries ({#1})}}

\newdimen \www 
\setbox0=\hbox{\n{viii}~}
\www=\wd0 \advance \www by .4em
\def\nnInContents#1#2{\hbox to \www {\hfill\n{#1}~\hskip .4em} #2\\}

\setbox0=\hbox{\sfdefault\large(00)}
\newdimen\howwide \howwide=\wd0 \howwide=2.5em

\def\sec#1#2#3{\setbox0=\hbox{\n{#1}~~}\leftskip \wd0
%\begin{flushleft}
\subsubsection*{\setbox0=\hbox{\n{#1}~~}\hskip -\wd0\copy0 \bfseries
\immediate\write\minicontentsfile{\bslash nnInContents{#1}{#3}}%
{#2}\starredsubsectioncontents{\hbox to \howwide{\hfill (#1)} #3}}
%\end{flushleft}
\leavevmode
\vskip 1ex
\par\leftskip 0in}

\def\highlightsec#1#2#3{\sec{#1}{{#2}}{{#3}}}
\def\highlightmysubsection#1{\subsection{{#1}}\subsectioncontents{{#1}}}
\def\changemysubsection#1{\subsection{{#1}}\subsectioncontents{{#1}}}
\def\changemysection#1{\section{{#1}}\sectioncontents{{#1}}}
\def\highlightmysection#1{\section{{#1}}\sectioncontents{{#1}}}
\def\mysection#1{\section{#1}\sectioncontents{#1}}
\def\mysubsection#1{\subsection{#1}\subsectioncontents{#1}}

\makeatletter
\let \bslash = \@backslashchar
\makeatother

\def\sectioncontents#1{\immediate\write\contentsfile{\bslash contents{\arabic{section} #1}{\thepage}}}
\def\subsectioncontents#1{\immediate\write\contentsfile{\bslash subcontents{\alph{subsection}\ #1}{\thepage}}}
\def\starredsubsectioncontents#1{\immediate\write\contentsfile{\bslash subcontents{#1}{\thepage}}}

\noindent\vbox{% The table of contents
\def\contents#1#2{\vskip 0mm \noindent{\bfseries #1} \dotfill\ #2\\}
\def\subcontents#1#2{\hbox to 1em{}{#1} \dotfill\ #2\\}
\noindent
% expand generated-supplementary.toc
\contents{1 Further issues for Software Engineering Boards (SEBs)}{3}
\subcontents{a\ {Brief definition}}{3}
\subcontents{b\ Relationships of SEBs to Ethics Boards}{4}
\subcontents{c\ {SEBs are necessary but not sufficient}}{5}
\contents{2 Software engineering best practice}{5}
\subcontents{a\ Introduction and standard references}{5}
\subcontents{b\ Essential components of best practice}{6}
\subcontents{\hbox to \howwide {\hfill (0)} Requirements}{6}
\subcontents{\hbox to \howwide {\hfill (1)} {Formal methods}}{6}
\subcontents{\hbox to \howwide {\hfill (2)} Defensive programming}{8}
\subcontents{\hbox to \howwide {\hfill (3)} {Using dependable programming languages}}{8}
\subcontents{\hbox to \howwide {\hfill (4)} Open source and version control}{8}
\subcontents{\hbox to \howwide {\hfill (5)} Rigorous testing}{10}
\subcontents{\hbox to \howwide {\hfill (6)} {Good documentation and record keeping}}{10}
\subcontents{\hbox to \howwide {\hfill (7)} Usability}{11}
\subcontents{\hbox to \howwide {\hfill (8)} Reusing quality solutions}{12}
\subcontents{\hbox to \howwide {\hfill (9)} Simplicity}{12}
\subcontents{\hbox to \howwide {\hfill (10)} {Compliance with standards}}{13}
\subcontents{\hbox to \howwide {\hfill (11)} Effective multidisciplinary teamwork}{13}
\subcontents{\hbox to \howwide {\hfill (12)} {Continuous Professional Development (CPD)}}{13}
\subcontents{\hbox to \howwide {\hfill (13)} {Security and other factors}}{14}
\contents{3 Code, data and publication}{14}
\contents{4 {The Speigelhalter trustworhiness questions}}{17}
\subcontents{a\ {How trustworthy are the numbers?}}{17}
\subcontents{b\ {How trustworthy is the source?}}{18}
\subcontents{c\ {How trustworthy is the interpretation?}}{18}
\contents{5 {Summary of pilot survey}}{21}
\subcontents{a\ {Assessment criteria and methods}}{21}
\subcontents{b\ {Detecting and defending against error}}{24}
\subcontents{c\ {Code policies of sampled journals}}{25}
\subcontents{d\ {Assessments and criteria}}{26}
\subcontents{e\ {Summary of assessments}}{30}
\subcontents{f\ {References for sampled papers}}{31}
% end expanding generated-supplementary.toc
}

\newwrite\contentsfile
\immediate\openout\contentsfile=generated-supplementary.toc

\newpage
\mysection{Further issues for Software Engineering Boards (SEBs)}

\highlightmysubsection{Brief definition}
Software Engineering Boards, henceforth SEBs, will be used to help and assure that critical code, including epidemic modeling, is of high standard, to provide assurance for scientific papers, Government public health and other policies, etc, that the code used is of appropriate quality for its intended uses.

Further details of the SEB proposal is in the main paper. Here we raise further issues for SEBs (additional to those covered in the main paper's introduction to SEBs), potential limitations and possible responses that can be addressed over time:

\begin{enumerate}\raggedright
\item 
Until there are national qualifications, nobody --- certainly nobody without professional training in software --- really knows just how bad (or good) they are at software engineering.

\item
When code is taken seriously, concerns may be raised on programmers' contributions to research, intellectual property rights, and co-authoring \cite{vancouver}. Software engineering is a hard, creative discipline, and getting epidemiological (and other scientific) models to work is generally a significant challenge, on a par with the setting up and exploring the mathematical models themselves. Often software engineers will need to explore boundary cases of models, and this typically involves hard technical mathematics \cite{hamming}. Often the software engineers will be solving entirely new problems and contributing to the research. How this is handled needs exploring. How software engineers are appropriately credited and cited for their contributions also needs exploring.

\item 
SEBs require policies on professional issues such as membership, transparency, and accountability.

\item
There should be a clear separation between the SEB members' activities as part of the Board, and their other activities, including professional advice, code development, or training (which is likely to be in demand from the same people who require formal approvals from the SEBs).

\item
Professional Engineering Bodies have a central role to play in professionalism, ranging from education and accreditation to providing professional structures and policies for SEBs. For example, should and if so how should the programming skills taught to computational scientists (epidemiologists, computational biologists, economists, computational chemists,~\ldots\@) be accredited?

\item
In the main paper, SEBs are viewed as a constructive contribution to good science, specifically helping improve the quality of epidemiological modeling. More generally, SEBs will have wider roles, for instance in overseeing software subject to medical device regulation \cite{fixit}.

\item
SEBs may fruitfully collaborate with other engineering disciplines to share and develop best practice. For example, engineers in other domains (e.g., civil engineers) routinely sign off projects, yet, on the other hand, they often overlook the quality of software engineering their projects implicitly rely on  --- for the same reasons as the scientific work discussed in this paper overlooks the dependence on quality software.

\item
Clearly, at least while this paper's concepts are tested and mature, SEBs will need to collaborate closely with research organizations, journals, and funding agencies in order to develop incremental developments to policies and processes that will be most effective, and which can be introduced most productively over time to the scientific community at large. Funding agencies may wish to support such strategic work, as they have previously funded one-off projects such as \cite{cosmos}.
\end{enumerate}

There are other ideas to help make SEBs work, but it is clear they are part of the solution. We must not let perfection be the enemy of the good. SEBs don't need to be perfect on day one, but they do need to get going in some shape or form to start making their vital contribution.

\mysubsection{Relationships of SEBs to Ethics Boards}

\begin{enumerate}\raggedright
\item 
Although SEBs may start with a checklist approach, like Ethics Boards generally do, it cannot be assumed that people approaching SEBs know enough about software engineering to perform adequate software assessments when there is any risk (as there is in public policy, medical apps, and so on). SEBs may also provide mentoring and training.

\item
Unlike Ethics Boards, which provide hands-off oversight, SEBs should provide professional advice, perhaps providing training or actually helping hands-on develop appropriately reliable software. During a pandemic SEBs would be very willing to do this, but in the long run it is not sustainable as voluntary labour, so all research, particularly medical research, should include support for professional software engineering.

\item 
Ethics Boards typically require researchers to fill in forms and provide details, which is a feasible approach as researchers know if they are doing experiments on children, for instance, so the forms are relatively easy to fill in (if often quite tedious). On the other hand, few healthcare and medical researchers understand software and programming, so they are \emph{not\/} able to fill in useful software forms on their own. SEBs need to know how well engineered the software really is, not how good its developers \emph{think\/} it is. As typical programs are enormous, SEBs are either going to need resources to evaluate programs, or they will need to supervise independent bodies that can do it for them.

\item
SEBs should have a two-way collaboration with Ethics Boards.

\begin{itemize}
\item SEBs have to deal with ethical concerns, and how they may be implemented in code. One of the papers \cite{ethics-paper} in the survey (discussed later in this \supplement) is a case in point, as is the growing cross-fertilization between AI and ethics \citeeg{ai-ethics}.

\item Ethics Boards also have to deal with software, and it is clear that they often fail to do this effectively. The case of the retraction of a peer reviewed articles for \emph{The Lancet\/} \cite{science-lancet1,science-lancet2,lancet-learning} and the \emph{Journal of Vascular Surgery\/} \cite{jvs1,jvs2,jvs3}, discussed in the main paper, are cases in point.
\end{itemize}

\item
Like some Ethics Boards, SEBs might become, or be perceived as becoming, onerous and heavy handed --- as if the Board is not interested in ethics but only in following a bureaucratic pathway. It seems essential, then, that SEBs have (and perhaps are chaired by) experienced, practicing, professional software engineers to avoid this problem. 
\end{enumerate}

\changemysubsection{SEBs are necessary but not sufficient}
The main paper provides evidence and argues that SEBs (or equivalent) are necessary to help improve the quality of science, specifically science relying, explicitly or implicitly, on tools or methods based in software.

SEBs address the problems identified at the laboratory end of doing science; they do not address the processes of review, editorial control, and action based on claimed results. As shown in the review of \plural{\dataN}{paper}, only some journals have code policies, and the policies are not enforced. In other words, improving the professionalization of software engineering has to proceed from doing science, which the paper covers, to the downstream issues of review and publication. SEBs may work with journals, funding agencies and even international standards agencies to improve broader awareness of professional software engineering, but this is a topic the present paper has not addressed. It needs doing.

\mysection{Software engineering best practice}
\label{supplementary-best-practice}
\mysubsection{Introduction and standard references}

This \supplement\ provides more explanations and justification for following standard software engineering practices that support reliable modeling, reliable research, and, most generally, reliable science.

The reader is referred to standard textbooks for more information \citeeg{sommerville,knight}, as well as to specialized texts that are more specifically addressed to software engineering in science \citeeg{cosmos}.

The book \emph{Why Programs Fail\/} \cite{wpf} is a very good practical guide to developing better code, and will be found very accessible. Humphrey \cite{humphrey} outlines a thorough discipline for anyone wanting to become a good programmer. Improvement is such an important activity, Humphrey has also published a book to persuade managers of the benefits \cite{managers}. Further suggestions for background reading can be found throughout this section.

\mysubsection{Essential components of best practice}
Software Engineering includes the following topics, which are discussed at more length below:
\vskip 2ex

\vbox{\noindent
      % expand generated-minicontents.tex
\nnInContents{0}{Requirements}
\nnInContents{1}{{Formal methods}}
\nnInContents{2}{Defensive programming}
\nnInContents{3}{{Using dependable programming languages}}
\nnInContents{4}{Open source and version control}
\nnInContents{5}{Rigorous testing}
\nnInContents{6}{{Good documentation and record keeping}}
\nnInContents{7}{Usability}
\nnInContents{8}{Reusing quality solutions}
\nnInContents{9}{Simplicity}
\nnInContents{10}{{Compliance with standards}}
\nnInContents{11}{Effective multidisciplinary teamwork}
\nnInContents{12}{{Continuous Professional Development (CPD)}}
\nnInContents{13}{{Security and other factors}}
% end expanding generated-minicontents.tex
      }

\newwrite\minicontentsfile
\immediate\openout\minicontentsfile=generated-minicontents.tex

\sec{0}{Without defining requirements, not enough skilled effort will be put into designing and implementing reliable software --- or excess effort will be wasted}{Requirements}
It is not always necessary to program well if the code to be produced is for fun, experimenting, or for demonstrations. On the other hand, if code is intended for life-critical applications, then it is worth putting more engineering effort into it. The first step of software engineering, then, is to assess the requirements, specifically the reliability requirements of the code that is going to be produced.

In practice, requirements and expectations change. Early experimental code, developed informally, may well be built on later to support models intended to inform public policy, for instance. Unfortunately, prototypes may impress project leaders who then want to rush into production software because, it seems, ``it obviously works.'' Fortunately, best practice software engineering can be adopted at any stage, particularly by using \emph{reverse engineering}. In reverse engineering, one carefully works out (generally partly automatically) what has already been implemented. This specification, carefully reviewed, is then used as the basis for a more rigorous software engineering process that implements a more reliable version of the system.

\highlightsec{1}{Without formal methods, there is no rigorous and checked specification of a program, so nobody --- including its developers --- will know exactly what it is supposed to do}{Formal methods}

{In the physical world, to do something as simple as design and build a barbecue, you would need to use elementary mathematics to calculate how many bricks to buy. To build something more substantial, such as block of flats, you would need to use structural engineering (with certified structural engineers) to ensure the building was safe. Although programming lends itself to mathematical analysis, it is surprising that few programmers use explicit mathematics at all in the design and implementation of software.}

{The type and use of mathematics used in software engineering is formal methods. Not using formal methods ensures the resulting code is unsafe and unreliable. Of particular relevance to scientific modeling: there must be an explicit use of formal methods to ensure  mathematical models (such as differential equations) are correctly implemented in code (and to understand the any limitations of doing so).}

Formal methods require sophisticated knowledge of logic \cite{cbc}, as well as practical knowledge of using appropriate formal methods tools (Alloy, HOL, PVS, SPARK, and \emph{many\/} others). Using the right tools is essential for reliable programming, because the tools do quickly and reliably what, done by hand, would be slow and error-prone. Standard tools cover verification, static analysis of code, version control, documentation, and so on --- this paper explains why some of these activities are essential for reliable programming below.

Crucially, tools are designed to catch common human errors that we are all prone to. Many tools are designed to avoid common human errors arising \emph{in the first place\/}; notably, the MISRA C toolset simply stops the developer using the most error-prone features of normal C, and hence improves the quality of programming with little effort.

>>>>

\begin{tabular}{lp{4.5in}}
\flagStyle{$\sf P_c$}&Journal has a code policy (see section~\ref{supplementary-journal-policies-section})\\
\flagStyle{$\sf P_{{\mbox{\scriptsize c-breach}}}$}&Paper breaches journal code policy (see section~\ref{supplementary-journal-policies-section})\\
\flagStyle{$\sf R_c$}&Paper uses a code repository (e.g., GitHub)\\
\flagStyle{$\sf R_{\mbox{\scriptsize c-empty}}$}&Code repository contains no code\\
\flagStyle{$\sf R_d$}&Paper uses a data repository (e.g., Dryad, Figshare, GitHub)\\
\flagStyle{$\sf S_{\mbox{\scriptsize NONE}}$}&No code available at all (note: code is not expected for standard models, systems or statistical methods)\\
\flagStyle{$\sf S_p$}&Paper says source code is available in principle\\
\flagStyle{$\sf S_{+}$}&Paper or URL provides source code\\
\flagStyle{$\sf S_{{\mbox{\scriptsize rigorous}}}$}&Evidence that source code was developed rigorously\\
\flagStyle{$\sf S_{\mbox{\scriptsize tested}}$}&Evidence that source code has been run with a clean build and tested\\
\flagStyle{$\sf S_{{\mbox{\scriptsize tools}}}$}&Evidence of any tool-based development\\
\flagStyle{$\sf S_{{\mbox{\scriptsize open source}}}$}&Team or open source development\\
\flagStyle{$\sf S_{{\mbox{\scriptsize otherSE}}}$}&Other evidence of good practice; see details in summary table\\
\flagStyle{$\sf C_0$}&Code has no non-trivial comments\\
\flagStyle{$\sf C_1$}&Code only has trivial comments (e.g., copyright)\\
\flagStyle{$\sf C_2$}&Helpful comments explaining code intent, rather than rephrasing the code\\
\flagStyle{$\sf C_{+}$}&Code has substantial, useful comments and documentation\\
\end{tabular}
